---
title: "Final Project"
author: "Mathias Heinz"
date: "2024-04-8"
categories: [code, analysis]

---
```{r}
library(tidyverse)
library(shiny)
library(plotly)

wine_ratings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-28/winemag-data-130k-v2.csv")

wine_clean <- wine_ratings |> filter(variety == "Cabernet Sauvignon" 
                                     | variety == "Pinot Noir" | variety == "Merlot" 
                                     | variety == "Chardonnay" | variety == "Sauvignon Blanc"
                                     | variety == "Pinot Gris" |  variety == "Reisling"                           | variety == "Red Blend"
                                     | variety == "White Blend" | variety == "Malbec"
                                     | variety == "Rosé" | variety == "Moscato"
                                     | variety == "Prosecco" | variety == "Tempranillo"
                                     | variety == "Syrah" | variety == "Shiraz"
                                     | variety == "Zinfandel") |> 
  mutate(r_w = case_when(variety == "Chardonnay" 
                          | variety == "Sauvignon Blanc" 
                          | variety == "Pinot Gris" 
                          | variety == "Reisling" 
                          | variety == "White Blend"
                          | variety == "Moscato"
                          | variety == "Rosé"
                          | variety == "Prosecco" ~ "White",
         variety == "Cabernet Sauvignon" | variety == "Pinot Noir"
         | variety == "Merlot" | variety == "Red Blend" | 
           variety == "Malbec" | variety == "Tempranillo" |
           variety == "Syrah" | variety == "Shiraz" |
           variety == "Zinfandel" ~ "Red"))
grape_sel <- wine_clean |> distinct(variety) |> pull(variety)
country_sel <- wine_clean |> distinct(country) |> pull(country)
wine_clean |> filter(!is.na(price)) |> summarise(min(price))
wine_clean |> filter(!is.na(price)) |> summarise(max(price))
## Of the 707 varieties in the data set I have collected 17 of the most popular grape varieties, leaving me with almost 60,000 observations.

```
### Written Report 

Abstract: 
  This report uses data collected by wine enthusiasts and submitted to git hub's tidytuesdays to analyze how wine production looks around the world. The data set contains wines from many countries around the world, with various information on each wine. Through my analysis I found the US to be have the highest number of wines in the world followed by France. Pinot Noir and Chardonnay had the highest variety of wines in both of these countries for red and white wines respectively.

Introduction: 
  For this project I wanted to explore wine production around the world. My overall goal was to look at some of the biggest wine producing regions, and see what type of wines they are producing, and the quality of these wines. The data set I started with had over 150,000 observations, on numerous countries and grape varieties, so I choose to filter the data set down to 17 varieties with a little under 60,000 observations. From this cleaned data set I focused on the variables "country", "variety", "points", "price", and sometimes "region". I also created the variable "r_w" to denote whether a wine is a red wine or a white wine. Through various graphs and the use of my shiny app I aimed to determine which countries in the world had the most wines, and how these wines were distributed among red and white wines, and grape variety. Next I wanted to know about the relationship between price and points rating for various wine varieties.


```{r}
wine_country <- wine_ratings |> mutate(country = fct_recode(country,
                                      "USA" = "US")) |>
  group_by(country) |> 
  summarise(total_wine = n())


world_df <- map_data("world")
world_full<- left_join(world_df, wine_country,
                       by = c("region" = "country"))
world_plot <- ggplot(data = world_full, aes(x = long, y = lat, group = group, 
                              label = region))+
  geom_polygon(aes(fill = total_wine ))+
  coord_map(projection = "mercator", xlim = c(-180,180))+
  scale_fill_viridis_c(direction = -1, option = "D")+
  labs(title = "Map of the World Colored by Total Number of Unique Wines")

ggplotly(world_plot, tooltip = "label")
```

  Here we can see a map showing the whole world. The countries colored in are countries included in the data set that I cleaned. Each country is colored by the total amount of observations for each country. More simply, counting up the total number of wines listed for each country. The United States has noticeably the highest amount coming in over 50,000 different wines. Following the US appears to be France and Italy who both fall in the 20,000 - 30,000 range. Some other notable countries include Argentina, Spain and Portugal who have slightly higher production than the other countries shown.

```{r}
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

world <- ne_countries(scale = "medium", returnclass = "sf")
Europe <- world[which(world$continent == "Europe"),]
europe_full<- left_join(Europe, wine_country,
                       by = c("name_sort" = "country"))

europe_plot <- ggplot(data = europe_full)+
  geom_sf(aes(fill = total_wine ))+
  coord_sf(xlim = c(-25,50), ylim = c(35,70), expand = FALSE)+
  scale_fill_viridis_c(direction = -1, option = "D")+
  labs(title = "Map of Europe Colored by Total Number of Unique Wines")
europe_plot

#Code for the outline of Europe retrieved from RPubs. Post by Misha Kollontai from 11-27-2019. URL: https://www.rpubs.com/mkollontai/DATA607_TidyVerse
```

  Since the map of the world makes it difficult to distinguish the countries of Europe, I decided to also map just the countries of Europe continuing to use the total_wine variable. This allows us to see the distinction between each country more clearly. We see again that France and Italy have the highest totals with France being slightly higher than Italy. Futher we see that Spain has a slightly higher number than Portugal, but both are very similar. However, this map allows us to see that Austria and Germany produce more distinct wines compared to other parts of western Europe.

## Italy Analysis
```{r}
wine_i <- wine_clean |> filter(country == "Italy") |> filter(!is.na(variety))
ggplot(wine_i, aes(x = price, y = points))+
  geom_hex()+
  facet_wrap(~r_w)+
  scale_color_viridis_c()+
  labs(title = "")

  
ggplot(wine_clean |> filter(country == "Italy") |> filter(r_w == "Red"), 
       aes(x = variety))+
  geom_bar(fill = "magenta4")+
  labs(title = "Red Wines From the Italy")

ggplot(wine_clean |> filter(country == "Italy") |> filter(r_w == "White"), 
       aes(x = variety))+
  geom_bar(fill = "yellow3")+
  labs(title = "White Wines From the Italy")
```
  
  The country of Italy appears to produce mainly Red Blend and White Blend wines. Italy produces over 300 different Red Blend varieties, and just under 800 varieties of White Blend wines. In the production of white wines, Prosecco (a sparkling white wine) and Moscato (a sweet white wine) are the next most produced, both around 200 varieties. The majority of Italian wines score between 85 and 90 points, while having a price per bottle around $25.

## France Anaylsis
```{r}
wine_f <- wine_clean |> filter(country == "France") |> filter(!is.na(variety))
ggplot(wine_f, aes(x = price, y = points))+
  geom_hex()+
  scale_color_viridis_c()

  
ggplot(wine_clean |> filter(country == "France") |> filter(r_w == "Red"), 
       aes(x = variety))+
  geom_bar(fill = "magenta4")+
  labs(title = "Red Wines From the France")
  
ggplot(wine_clean |> filter(country == "France") |> filter(r_w == "White"), 
       aes(x = variety))+
  geom_bar(fill = "yellow3")+
  labs(title = "White Wines From the France")

france_cs <- wine_clean |> filter(country == "France") |> 
  filter(variety == "Cabernet Sauvignon")

france_plot <- ggplot(data = france_cs, aes(x = price, y = points, label = title))+
  geom_point()+
  geom_jitter(width = .05)
  theme_minimal()+
  labs(title = "Price vs Points for French Cabernet Sauvignon")

ggplotly(france_plot, tooltip = "label")
```
  
  From my Europe map I found France to be the largest wine producing country in Europe. Diving a little deeper into the data, I found that Pinot Noir is the most produced red variety, just under 2000 different bottles. For French white wine, Chardonnay was the most produced at a little over 2500 bottles. I found the prevalence of Rosé to be a bit surprising, as it was the second most common variety in the white wine group. This could be explained by the recent rise in popularity of Rosé.
  
## US Analysis
```{r}
ggplot(wine_clean |> filter(country == "US") |> filter(r_w == "Red"), 
       aes(x = variety))+
  geom_bar(fill = "magenta4")+
  labs(title = "Red Wines From the United States")
  
ggplot(wine_clean |> filter(country == "US") |> filter(r_w == "White"), 
       aes(x = variety))+
  geom_bar(fill = "yellow3")+
  labs(title = "White Wines From the United States")

wine_u <- wine_clean |> filter(country == "US") |> filter(!is.na(variety))
ggplot(wine_u, aes(x = price, y = points))+
  geom_hex()+
  scale_color_viridis_c()+
  labs(title = "Hex Plot of price vs points for United States Wines")

```
  
  As I found the US to have the largest number of wines, I wanted to see how the distribution looked between red and white wines, as well as see the the distribution of US wines based on price vs points. I found that Pinot Noir was the most common red variety with almost 10,000 observations. Following Pinot Noir is Cabernet Sauvignon with just under 7,500 different wines. The hex plot shows that the highest density of US wines are between $1-$125 and score around 90 points.

```{r}
cali_cs_pn <- wine_clean |> filter(country == "US") |> 
  filter(variety == "Cabernet Sauvignon" | variety == "Pinot Noir") |> 
  filter(province == "California")
wine_cali <- wine_clean |> filter(country == "US") |> 
  filter(province == "California")

ggplot(wine_cali |> filter(r_w == "Red"), aes(x = variety))+
  geom_bar(fill = "magenta4")+
  labs(title = "Red Wines of California, US")

ggplot(cali_cs_pn, aes(x = variety, y = price))+
  geom_boxplot(fill = "magenta4", color = "yellow3")+
  labs(title = "Boxplot of Price for Cabernet Sauvignon and Pinot Noir from California")

```

  One of the most famous wine growing regions in the United States is California, and its wineries are known for their Cabernet Sauvigon's. Despite this, in the bar chart we can see that there are actually more individual bottles of Pinot Noir, but Cabernet Sauvignon is a close second. Next, the boxplot shows that although there are more Pinor Noirs, Cabernet Sauvignon is slightly more expensive on average, both around $50. However, there are no Pinor Noir bottles over $200, while Cabernet Sauvignon has many bottles over $200 and one more than $600.
  
  
## New Zealand Analysis
```{r}
wine_n <- wine_clean |> filter(country == "New Zealand") |> filter(!is.na(variety))


ggplot(wine_clean |> filter(country == "New Zealand") |> filter(r_w == "White"), 
       aes(x = variety))+
  geom_bar(fill = "yellow3")+
  labs(title = "White Wines From the New Zealand")

newz_sb <- wine_clean |> filter(country == "New Zealand") |> 
  filter(variety == "Sauvignon Blanc")
nz_plot <- ggplot(newz_sb, aes(x = price, y = points, label = title))+
  geom_point()+
  scale_color_viridis_c()+
  labs(title = "Price vs Points for New Zeland Sauvignon Blanc")
ggplotly(nz_plot, tooltip = "label")
```

  New Zealand is a country known for their white wine and specifically their Sauvignon Blanc. This can be seen in the bar graph, as Sauvignon Blanc dwarfs all other white wines. I was then curious about the ratings and price points of these Sauvignon Blancs. My scatter plot revealed that Brancott 2013 Chosen Rows Sauvginon Blanc (Marlborough) is both the highest rated and most expensive Sauvignon Blacn, with a score of 94 and a price around $75. The graph also shows that most New Zealand Sauvignon Blanc's are around $20 with a score between 85 and 90. 

## Future Work:

If I had more time I this projet I would have loved to continue to tweak my shiny app. For example I would have liked to let the user choose between either hex or scatter plot depending on the total number of observations. Further I would have liked to add more specific location options like region or state. Lastly if I would have gotten the color scheme to change based on the red or white wine input that would have looked nice. I would also like to acknowledge some limitations of my app. First, I have to mention that the data set I created does not contain all wines in the world or all the grape varieties. For the sake of the project I thought it best to only include 17 of the most well known grape varieties in my analysis and app. This fact also allows the app to be more user friendly for a wine novice.  

## Shiny App: 

The following is the shiny app I created to aid in my analysis, as well as a nice tool to examine wine in specific countries. This app allows the user to select a country, either red or white wine, specific grape variety, and lastly a price range. Then the app produces three graphs and one table based on the user's chosen inputs. First a hex plot representing the chose wine is presented, showing price vs points. Second, a scatter plot again showing price vs points, that allows the user to hover over a point and see the specific name, price, and point score for that wine. Next is the table output for the users selection. This table shows all the information in the data set allowing the users to learn more about a wine that might catch their interest. In the table is more in depth information on the wines, including a description of the flavor and more precise location of origin within the chosen country.


```{r}
ui <- navbarPage("Wine of the World",
  sidebarLayout(
    sidebarPanel(
      selectInput("country_sel", label = "Choose a Country", 
                                 choices = country_sel),
      radioButtons("white_or_red", label = "Red or White",
                                  choices = c("Red", "White")),
      selectInput("grape_sel", label = "Choose a Grape",
                                 choices = grape_sel),
      sliderInput("price_range", label = "Choose a Minimum Price",
                                 min = 4, max = 2500, 
                  value = c(4,2500), step = 1)
    ),
    mainPanel(
      tabsetPanel(
        tabPanel("Hex Plot", plotOutput("wine_plot")),
        tabPanel("Wine Point", plotlyOutput("wine_point")),
  tabPanel("Wine Table", tableOutput("wine_table")),
  tabPanel("Bar Graph", plotOutput("wine_bar"))
      )
    )
  ))

server <- function(input, output, session) {
  observeEvent(input$white_or_red, {
   grape_sel <- wine_clean |> filter(r_w == input$white_or_red) |> distinct(variety) |> pull(variety)
   
   updateSelectInput(inputId = "grape_sel",
                     choices = grape_sel)
 })
  observeEvent(input$country_sel,{
    grape_sel <- wine_clean |> filter(r_w == input$white_or_red) |> filter(country == input$country_sel) |> distinct(variety) |> pull(variety)
    updateSelectInput(inputId = "grape_sel",
                      choices = grape_sel)
  })
   wine_df <- reactive({
    wine_one <- wine_clean |> filter(country == input$country_sel) |> 
  filter(variety == input$grape_sel) |> 
  filter(price >= input$price_range[1] & price <= input$price_range[2])
  })
   wine_bar <- reactive({
     wine_clean |> filter(country == input$country_sel) |> 
     filter(r_w == input$white_or_red)
   })
   output$wine_bar <- renderPlot({
    ggplot(wine_bar(), aes(x = variety))+
  geom_bar()+
  facet_wrap(~r_w)+
       coord_flip()
   })
  output$wine_table <- renderTable({
    wine_df()
  })
  output$wine_plot <- renderPlot({
    ggplot(wine_df(), aes(x = price, y = points))+
  geom_hex()+
  scale_color_viridis_c()+
  labs(title = glue::glue("Graph of Price vs Score for ", input$grape_sel, 
                          " Wines from ", input$country_sel ))
  })
  output$wine_point <- renderPlotly({
    ggplot(data = wine_df(), aes(x = price, y = points, label = title))+
      geom_point()+
      theme_minimal()+
      labs(title = glue::glue("Graph of Price vs Score for ", 
                              input$grape_sel, " Wines from ", input$country_sel ))
    
  })
}

shinyApp(ui, server)
```

## Link to Blog Post: https://mhein804.github.io/ds334blog/


